---
title: "Cyclistic Analysis - Google Data Analytics Capstone Project"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(tidyverse)
```

**Author**: Gisanrin Oluwaseun  
**Date Updated**: 20th February, 2024

**Notice**  
This publication was completed as a part of the Google Data Analytics course on Coursera. You may get in touch with me via email at [seungisanrin@gmail.com](mailto:www.seungisanrin@gmail.com).  

## Background
Cyclistic, a bike-share program launched in 2016 in Chicago, has grown significantly, boasting a fleet of 5,824 bicycles across 692 stations, offering geotracked and flexible rental options for 
customers. While its marketing strategy initially targeted broad consumer segments, recent financial analysis indicates that annual members generate higher profits compared to casual riders. 
Recognizing this, Moreno, Cyclistic's strategist, aims to shift focus towards converting casual riders into annual members. To achieve this, her team plans to delve into Cyclistic's historical bike 
trip data to discern trends and understand the differing behaviors of casual riders and annual members, ultimately informing targeted marketing strategies that leverage digital media effectively.

### Business Task
How do annual members and casual riders use Cyclistic bikes differently?

## Analysis Process
This details the procedure by which this analysis was completed.

### Data Preparation and Pre-processing
The dataset utilised for the completion of this task covered the span of one year from February 2023 to January 2024, the dataset can be downloaded [here](https://divvy-tripdata.s3.amazonaws.com/index.html).
The data has been made available by Motivate International Inc. under this [license](https://ride.divvybikes.com/data-license-agreement). It is a public dataset, and it can be used in exploring the differences
in the customer types for Cyclistic. The data is hosted on the Amazon Web Server (AWS), and is first-party data owned by Cyclistic.

#### Preparation
The language used in the preparation and cleaning of the dataset to be used for analysis is R. R is a dynamic programming language that can be used in data preparation, analysis, visualisations, as well as for
creating reports.  
The 12 datasets (monthly) used for this analysis, each with 13 columns and between 100,000 and 700,000+ rows. The datasets were combined into `origin_cyc_df` dataframe in R.
```{r dataset compilation, include=FALSE}
setwd("C:/Users/seung/OneDrive/Documents/GitHub/cyclistic-analysis/cyclistic-analysis/data")
origin_cyc_df <- list.files(path = "C:/Users/seung/OneDrive/Documents/GitHub/cyclistic-analysis/cyclistic-analysis/data") %>%
  lapply(read_csv) %>%
  bind_rows()
```
The combined dataframe has **5,674,449 rows** and **13 columns**.
```{r column details}
glimpse(origin_cyc_df)
```

#### Cleaning and Manipulation
A brief outline providing details on the steps taken to clean the dataset for analysis.  
+ The columns relevant for the analysis were selected, this includes (ride_id, rideable_type, started_at, start_station_name, ended_at, end_station_name, member_casual).  
+ Duplicates in the `ride_id` columns were checked for, and the column was determined to be unique.  
+ Checked the `ride_id` column for nonconforming observations, all observations are made up of 16 character strings. All observations of the `ride_id` variable were valid.  
+ Filtered through all columns to remove ride observations that had any of its variables blank.  
+ Created column `ride_length` to determine the length of each ride observation in minutes.  
+ Dropped observations with less `ride_length` less than 1 minute, which are potentially false starts or attempts at re-docking bikes. As well as observations with negative values for `ride_length`.  
+ Created column `weekday` to denote the day of the week a ride was embarked on.  
  
The cleaning of the dataset led to the removal of **1,464,313** ride observations.  
The dataset for analysis is composed of **4,210,136** rows and **9** columns.

#### Analysis
For this step, summary statistics were performed to determine differences between members and casuals
riders. Number of valid rides were computed, the average length ride between the period of Feb-2023 and
Jan-2024, also the maximum ride length for the time period.  
Member-type analyses were conducted to determine the ride distribution among the 4 million rides, the 
frequency of trips on each day of the week, the average ride length, also the bike preferences for each
type of member.  
The detailed analysis and visualisations, can be found here.
